\protect A Web of Human Knowledge: Unravelling Wikipedia's First Link Network

### 2015-11-23 writeup/2015-11-wikipedia-network.tex
ndegree section updated

\section*{Introduction}

Wikipedia is a towering achievement of the modern era. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{stats} \cite{drugs}.
All this information does not lie dormant. 
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views in a single month. 
The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.

@@

\section*{Traversing the First Link Network}

@@

To more accurately gauge the distribution, we construct a log-log plot of the entire dataset: log(traversal visits) against log(rank). We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93. A handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

@@

Measuring only traversal visits however is limiting as it does not distinguish whether a particular article in a cycle 
is a funnel, directing many more paths inside the cycle than others. 
To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
so that accumulation does not flow to all articles in a cycle.
The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
from an article directly funneling many paths.

@@

@@

\section*{Appendix}

\subsection*{Constructing The First Link Network}

@@

\subsection*{Supplementary Tables}

@@

\end{thebibliography}

>>>>

\section{Introduction}

Wikipedia is a towering achievement of the modern era. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{stats} \cite{drugs}.
All this information does not lie dormant. 
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views in a single month
\cite{wiki_views}.
The most widely used general reference, Wikipedia has amassed an awe-inspiring collection of human knowledge.

@@

\section{Traversing the First Link Network}

@@

To more accurately gauge the distribution, we construct a log-log plot of the entire dataset: log(traversal visits) against log(rank). We observe a fairly linear fit, as is characteristic of scale-free networks, with an r-value of -.93 and 
a power-law exponent of $-0.636$. A handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

@@

\subsection{Degree Distribution}

We identified the set of first links directly referencing each article. 
Similar to the number of visits, we can examine the overall distribution and 
first links to understand the direct structure of taking one step through the 
FLN.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by n-degree}
  }
  We rank each article by the number of direct first links to the article (n-degree). 
  \label{fig:ndegree list}

\end{figure*}
The highest ranking article by n-degree, or the number of direct first links, is 
"United States" with 80,249 direct first links. Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands."

"Philosophy" and other Philosophical concepts which hold many traversal visits,
do not appear near the top when ranked by n-degree. 
"Philosophy" has an n-degree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy," the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of first links, as we argued in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears scale free with most articles having fewer than 9 
  direct first links, while few hold most.
  \label{fig:degree distribution}

\end{figure*}
As a distribution, we examine the degree of the articles in the FLN with at least
one direct first link. The average n-degree 
is 3.6 direct links with a standard deviation of 89.5 links.
Only 4826 articles have more than 100 direct first links, $75\%$ have fewer than 
9. Overall the n-degree appears scale-free with an r value of -0.98 as a linear log-log fit and a power law exponent of -0.79.

@@

Measuring only traversal visits and n-degree is limiting as it does not distinguish whether a particular article in a cycle 
is a funnel, directing many more paths inside the cycle than others. 
To distinguish among articles in a cycle, we also measured traversal funnels, or the number of 
paths an article directs towards a cycle. Here we count the number of traversal visits up to a cycle, 
so that accumulation does not flow to all articles in a cycle.
The importance here is to distinguish an article that happens to be connected to another article with many traversal visits 
from an article directly funneling many paths.

@@

\red{TODO: add two regimes for power-law distribution}

\subsection{Article Popularity}

Using a recently released api, we compare the popularity of a page to our 
findings about the FLN. We measure popularity as the total number 
of page views in the english edition of Wikipedia in the month of 
October 2015 (the earliest full month for which the data is available). 
We find \dots

@@

\section{Appendix}

\subsection{Constructing The First Link Network}

@@

\subsection{Supplementary Tables}

@@

    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}
\end{thebibliography}
